An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2017-03-30 uai2017/paper.tex
Final touch-ups. Will send to arxiv, UAI

Our test is applicable when the variance (over data samples) of the
log probability ratio between proposal and current state is small
enough (less than 1). It's not clear at first why this quantity should
be bounded, but it is natural for well-specified models running
Metropolis-Hastings sampling with optimal
proposals~\citep{OptimalScaling01} on a full dataset. If the posterior
parameter distribution is a unit-variance normal distribution, then
the posterior for $N$ samples will have variance $1/N$. There is
simply not enough information in $M << N$ samples to locate and
efficiently sample from this posterior. This is not a property of any
particular proposal or test, but of the information carried by the
data. The variance condition succinctly captures the condition that
the minibatch carries enough information to generate a sample.  While
we cannot expect to generate independent samples from the posterior
using only a small subset of the data, there are three situations
where we can exploit small minibatches:

@@

We note that 2 above is characteristic of Gibbs samplers applied to
large datasets~\citep{dupuy2016}. Such samplers represent a
model posterior via counts over an entire dataset of $N$ samples. When
a minibatch of $M$ samples is used to update the model, the counts for
these samples only are updated. This creates ``steps'' of $O(M/N)$
in the model parameters, and correlated samples from the model
posterior. Correlated samples are still very useful in high-dimensional
ML problems with multi-model posteriors since they correspond to a finer-scale
random walk through the posterior landscape. 
\begin{comment}
It is worth discussing at this point the typical goals of M-H sampling on large
datasets. By the Bernstein-von Mises Theorem, the posterior distribution for a
Bayesian inference task has variance that scales inversely with $N$. Simply
sampling from it is one application, but an efficient
proposal~\citep{OptimalScaling01} has similar variance to the target and will
diffuse to it extremely slowly. For applications to neural networks or models
where the posterior is multimodal~\citep{choromanska2014loss}, samplers will
likely get trapped in one of the modes. A common solution is to anneal the
sampler, running first at high temperatures to flatten the likelihood landscape.
This in turn reduces the variance of the log acceptance probability and allows
our test to be applied. These samples can cover the search space densely with
small steps rather than taking a few sparse steps towards an optimum. In this
mode, M-H tests can be used in similar fashion to Stochastic Gradient Descent.
The goal in SGD is to make gradual progress to a posterior mode with each step,
taking small steps so that the cumulative displacement has progressively lower
variance.
\end{comment}
The contributions of this paper are as follows:

@@

    \item We compare our new test and prior approaches on several
      datasets. We demonstrate several order-of-magnitude improvements in sample efficiency,
      and that batch size distribution is short-tailed. 
\end{itemize}

@@

Table~\ref{tab:logistic_settings} indicates the relevant parameter settings for
the logistic regression experiments. Unless otherwise stated, values apply to
all methods tested. For values from~\citep{cutting_mh_2014}
or~\cite{icml2014c1_bardenet14}, we use their notation to be consistent.

>>>>

Our test is applicable when the variance (over data samples) of the
log probability ratio between the proposal and the current state is small
enough (less than 1). It's not clear at first why this quantity should
be bounded, but it is natural for well-specified models running
Metropolis-Hastings sampling with optimal
proposals~\citep{OptimalScaling01} on a full dataset. If the posterior
parameter distribution is a unit-variance normal distribution, then
the posterior for $N$ samples will have variance $1/N$. There is
simply not enough information in $M \ll N$ samples to locate and
efficiently sample from this posterior. This is not a property of any
particular proposal or test, but of the information carried by the
data. The variance condition succinctly captures the condition that
the minibatch carries enough information to generate a sample.  While
we cannot expect to generate independent samples from the posterior
using only a small subset of the data, there are three situations
where we can exploit small minibatches:

@@

We note that case two above is characteristic of Gibbs samplers applied to
large datasets~\citep{dupuy2016}. Such samplers represent a
model posterior via counts over an entire dataset of $N$ samples. When
a minibatch of $M$ samples is used to update the model, the counts for
these samples only are updated. This creates ``steps'' of $O(M/N)$
in the model parameters, and correlated samples from the model
posterior. Correlated samples are still very useful in high-dimensional
ML problems with multi-modal posteriors since they correspond to a finer-scale
random walk through the posterior landscape.  The contributions of this paper
are as follows:

@@

    \item We compare our new test and prior approaches on several
      datasets. We demonstrate several order-of-magnitude improvements in sample efficiency,
      and that the batch size distribution is short-tailed. 
\end{itemize}

@@

Table~\ref{tab:logistic_settings} indicates the relevant parameter settings for
the logistic regression experiments. Unless otherwise stated, values apply to
all methods tested. For values from~\citep{cutting_mh_2014}
or~\cite{icml2014c1_bardenet14}, we use their notation ($\Delta^*, m, \epsilon,
\gamma, p$, and $\delta$) to be consistent.
